---
title: "Lab 6.1 - Visualizing Spatial Data"
author: "Sherise Immanuela"
date: "Fall 2022"
output:
  html_document:
    code_download: true
---

```{r setup, include = FALSE}
# Set some useful chunk options for code chunks.
knitr::opts_chunk$set(
  echo = TRUE,
  error = TRUE,    # display errors but keep going.
  comment = "",    # don't add '##' characters before output.
  message = FALSE  # don't include messages in the knitted output (check them in RStudio)
  )
library(tidyverse)
theme_set(theme_bw())
```

The late comedian [Mitch Hedberg](https://en.wikipedia.org/wiki/Mitch_Hedberg) famously joked that "La Quinta" is Spanish for "next to Denny's". In this lab exercise, we explore the accuracy of his claim that [Denny's](https://www.dennys.com/), a casual diner chain that is open 24 hours, is actually co-located with [La Quinta Inn and Suites](http://www.lq.com/), a hotel chain. This lab is inspired by a blog post by John Reiser entitled [Mitch Hedberg and GIS](http://njgeo.org/2014/01/30/mitch-hedberg-and-gis/).

## Loading the Dataset

We use three datasets, all stored in [./data](https://cs.calvin.edu/courses/info/601/06wrangling/lab/data/.):

- `dennys` --- The locations of all Denny's restaurants:
  - *Scraped* from [All Denny's Locations](https://locations.dennys.com/)
  - Described here: [Locations of Denny's restaurants](https://rstudio-education.github.io/dsbox/reference/dennys.html)
  
- `laquinta` --- The locations of all La Quinta hotels: 
  - *Scraped* from [La Quinta Locations](https://www.wyndhamhotels.com/laquinta/locations)
  - Described here: [Locations of La Quinta Inn motels](https://rstudio-education.github.io/dsbox/reference/laquinta.html)

- `states` --- The names of all the US states, including DC, plus some additional information.

Note that the first two datasets are represented in RDS format, which is an R-internal format for datasets that maintains data types. The last dataset is in CSV format and requires some 

```{r load-data, message=FALSE}
dennys <- read_rds("data/dennys.rds")
laquinta <- read_rds("data/laquinta.rds")
states <- read_csv("data/states.csv", col_types = cols(
  name = col_character(),
  abbreviation = col_character(),
  area = col_double()
))
```

Data Dictionary:

dennys (a tibble with 1643 rows and 6 variables)
`address`: Street address of the restaurant 
`city`: City of the restaurant.
`state`: US State of the restaurant.
`zip`: US ZIP Code of the restaurant.
`longitude`: The longitude of the restaurant in degrees.
`latitude`: The latitude of the restaurant. in degrees.

laquinta (a tibble with 909 rows and 6 variables)
`address`: Street address of the motel. 
`city`: City of the motel.
`state`: US State of the motel.
`zip`: US ZIP Code of the motel.
`longitude`: The longitude of the motel in degrees.
`latitude`: The latitude of the motel in degrees.

states (a tibble with 51 rows and 3 variables)
`name`: Name of the state.
`abbreviation`: Abbreviation of the state name.
`area`: Area of state.

## Focusing on US Locations

We limit our analysis to Denny"s and La Quinta locations in the United States.

```{r}
dennys %>%
  filter(!(state %in% states$abbreviation))
```
There are no Denny's locations outside the US.

```{r}
laquinta <- laquinta %>%
  filter((state %in% states$abbreviation))
```
There are 14 La Quinta records that were filtered out because they're not in the US.

## Computing Frequencies

We now compute some density statistics for the two franchises.

### By State

```{r}
dennysCount <- dennys %>% 
  group_by(state) %>%
  count(state, sort=TRUE)
dennysCount
```
California has the most Denny's locations (403), while Delaware has the least (1).

```{r}
laquintaCount <- laquinta %>% 
  group_by(state) %>%
  count(state, sort=TRUE)
laquintaCount
```
Texas has the most La Quinta locations (237), while Maine has the least (1).

### By Area

```{r}
dennysArea <- left_join(dennysCount, states, by = c("state" = "abbreviation"))

dennysDensity <- dennysArea %>%
  mutate(density = (n/area)*1000) %>%
  arrange(desc(density))
dennysDensity
```

The District of Columbia has the highest density by far, but mainly because the area is so small.
It's interesting that Texas does not have a high density, especially because it's the state with the 2nd highest number of Denny's.

```{r}
laquintaArea <- left_join(laquintaCount, states, by = c("state" = "abbreviation"))

laquintaDensity <- laquintaArea %>%
  mutate(density = (n/area)*1000) %>%
  arrange(desc(density))
laquintaDensity

```

Rhode Island has the highest density, which is mainly because the area is so small. 
Similar to the density of Denny's locations, Texas does not have a high density, even if the state has the most number of La Quinta locations by far.

## Plotting Locations

Finally, we plot the locations (latitude and longitude) of both establishments.

To do this, we'll put the two datasets together into a single data frame.
First, we'll add an identifier variable called `establishment` to distinguish rows from the two datasets, and then we'll bind them with the `bind_rows` function (n.b., this binding works because the two tables have the same columns):

```{r}
dn_lq <- bind_rows(
  dennys %>% mutate(establishment = "Denny's"), 
  laquinta %>% mutate(establishment = "La Quinta"), 
)
```

```{r}
dn_lq %>% 
  ggplot() +
  aes(x = longitude, y = latitude, color = establishment, alpha=0.5) +
  geom_point()
```

## Drawing Conclusions

In conclusion, Denny's restaurants and La Quinta motels appear to be visually co-located.
This is especially true for locations in the East Coast and West Coast, and even Midwest. 
