---
title: "DATA 202 Homework 6"
author: "Sherise Immanuela"
date: "`r Sys.Date()`"
output:
  html_document:
    code_download: true
---

```{r setup, include=FALSE}
# Set some useful chunk options for code chunks.
knitr::opts_chunk$set(
  echo = TRUE,
  error = TRUE,    # display errors but keep going.
  comment = "",    # don't add '##' characters before output.
  message = FALSE  # don't include messages in the knitted output (check them in RStudio)
  )
library(tidyverse)
library(tidymodels)
theme_set(theme_bw())
```

## Problem 1: Setting up the problem

a. Features incldue: building types, roof styles, conditions, etc.
b. Target: Find the sale price of a home
c. This is a regression task because we're predicting the price of a house sale.
d. One appropriate error metric for this task would be the mean absolute percent error.
e. This task is expected to produce a mean absolute percent error of 10%. 
f. An example of an inappropriate error metric is a mean absolute error. The mean absolute error is measured as the average of absolute error values, and it would be inappropriate because the range of values for the data varies widely.

## Problem 2: Fitting and evaluating a model
```{r}
ames_home_sales <- read_builtin("ames", package = "modeldata") %>% 
  mutate(Sale_Price = Sale_Price / 1000) %>% 
  filter(Gr_Liv_Area < 4000, Sale_Condition == "Normal")
```

a.
```{r}
set.seed(456)
ames_split <- initial_split(ames_home_sales, prop = 0.9)
ames_train <- training(ames_split)
ames_test <- testing(ames_split)
```

b.
```{r}
linear_reg_fit <- fit(linear_reg(mode = "regression"),
  Sale_Price ~ Gr_Liv_Area,
  data = ames_train)
linear_reg_fit
```

```{r}
ames_train_predictions <- linear_reg_fit %>% 
  augment(ames_train)
```

c.
```{r}
metrics <- yardstick::metric_set(rsq_trad, mae, mape, rmse)
ames_train_predictions %>% 
  metrics(truth = Sale_Price, estimate = .pred) %>% 
  select(-.estimator)
```

The mean absolute difference between the actual and predicted value (mae) is 33.4, which is about a 20% difference (mape). The percentage is not ideal, but acceptable.

d. The model is expected to predict results with a 20% difference than the actual value.

## Problem 3 (Repeat problem 1)
a. Features incldue: building types, roof styles, conditions, etc.
b. Target: Find whether a home will sell for above or below the median price of all homes in the dataset.
c. This is a regression task because we're predicting at what price a home will sell.
d. One appropriate error metric for this task would be the mean absolute percent error.
e. This task is expected to produce a mean absolute percent error of 10%. 
f. An example of an inappropriate error metric is a mean absolute error. The mean absolute error is measured as the average of absolute error values, and it would be inappropriate because the range of values for the data varies widely.

```{r}
ames_vs_median <- ames_home_sales %>% 
  mutate(sale_category = case_when(
    Sale_Price > median(Sale_Price) ~ "Above Median",
    TRUE ~ "Below Median") %>% 
    as_factor() %>% 
    fct_relevel("Above Median"))
```

```{r}
set.seed(123)
ames_vs_median_split <- initial_split(ames_vs_median, prop = 0.9) 
ames_vs_median_train <- training(ames_vs_median_split)
ames_vs_median_test <- testing(ames_vs_median_split)
```

```{r}
model <- decision_tree(mode = "classification") %>%
  fit(sale_category ~ Gr_Liv_Area, data = ames_vs_median_train)
model3
```

```{r}
model %>%
  extract_fit_engine() %>%
  rpart.plot::rpart.plot(roundint = FALSE)
```

```{r}
ames_vs_median_train_predictions <- model %>% 
  augment(ames_vs_median_train)
```

## Problem 4 (Repeat problem 2)

```{r}
levels(ames_vs_median$sale_category)
```

```{r}
metrics <- yardstick::metric_set(accuracy, sensitivity, specificity)
ames_vs_median_train_predictions %>% 
  metrics(truth = sale_category, estimate = .pred_class) %>% 
  select(-.estimator)
```

The accuracy metric shows that the model was correct 79% of the time. 
The sensitivity metric shows that the true positive rate of the model is 82%.
Given these metrics, the model can generally be accepted to predict whether a home will be sold for above or below the median home sales price.

