---
title: "DATA 202 Homework 3"
author: "Sherise Immanuela"
date: "`r Sys.Date()`"
output:
  html_document:
    code_download: true
---

```{r setup, include=FALSE}
# Set some useful chunk options for code chunks.
knitr::opts_chunk$set(
  echo = TRUE,
  error = TRUE,    # display errors but keep going.
  comment = "",    # don't add '##' characters before output.
  message = FALSE  # don't include messages in the knitted output (check them in RStudio)
  )
library(tidyverse)
theme_set(theme_bw())
```

## Loading Data
```{r}
dennys <- read_rds("data/dennys.rds")
laquinta <- read_rds("data/laquinta.rds")
states <- read_csv("data/states.csv", col_types = cols(
  name = col_character(),
  abbreviation = col_character(),
  area = col_double()
))
```

## Exercises
```{r}
dn_ak <- dennys %>%
  filter(state == "AK")
```

1. There are 3 Denny’s locations in Alaska. 

```{r}
lq_ak <- laquinta %>%
  filter(state == "AK")
```

2. There are 2 La Quinta locations in Alaska.

```{r}
dn_lq_ak <- full_join(dn_ak, lq_ak, by = "state",
                      suffix = c("_dn", "_lq"))
```

3. There are 6 pairings between all Denny’s and all La Quinta locations in Alaska.
4. There are 6 observations in the joined dn_lq_ak data frame. 

```{r}
# Great-circle distance, Implementation from dsbox
haversine <- function(long1, lat1, long2, lat2) {
  # convert to radians
  long1 = long1 * pi / 180
  lat1  = lat1  * pi / 180
  long2 = long2 * pi / 180
  lat2  = lat2  * pi / 180
  
  # Earth mean radius in km (WGS84 ellipsoid)
  R = 6371.009
  
  # Compute the distance in km
  a = sin((lat2 - lat1)/2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1)/2)^2
  d = R * 2 * asin(sqrt(a))
  return(d)
}

dn_lq_ak <- dn_lq_ak %>%
  mutate(distance = haversine(longitude_dn, latitude_dn, longitude_lq, latitude_lq))
```

```{r}
nearest_dn_lq <- dn_lq_ak %>% 
  group_by(address_lq) %>%
  count(distance, sort=TRUE) %>%
  slice_head()
```

There are 2 La Quinta locations in Alaska that are “next to Denny’s”.

## Exercises - Maryland
```{r}
dn_md <- dennys %>%
  filter(state == "MD")
```

```{r}
lq_md <- laquinta %>%
  filter(state == "MD")
```

```{r}
dn_lq_md <- full_join(dn_md, lq_md, by = "state",
                      suffix = c("_dn", "_lq"))
```

```{r}
# Great-circle distance, Implementation from dsbox
haversine <- function(long1, lat1, long2, lat2) {
  # convert to radians
  long1 = long1 * pi / 180
  lat1  = lat1  * pi / 180
  long2 = long2 * pi / 180
  lat2  = lat2  * pi / 180
  
  # Earth mean radius in km (WGS84 ellipsoid)
  R = 6371.009
  
  # Compute the distance in km
  a = sin((lat2 - lat1)/2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1)/2)^2
  d = R * 2 * asin(sqrt(a))
  return(d)
}

dn_lq_md <- dn_lq_md %>%
  mutate(distance = haversine(longitude_dn, latitude_dn, longitude_lq, latitude_lq))
```

```{r}
nearest_dn_md <- dn_lq_md %>% 
  group_by(address_lq) %>%
  count(distance, sort=TRUE) %>%
  slice_head()
```

There are 6 La Quinta locations in Maryland that are less than 6 miles from the nearest Denny's. 