---
title: "DATA 202 Homework 4"
author: "Felicia Susanto"
date: "`r Sys.Date()`"
output:
  html_document:
    code_download: true
---

```{r setup, include=FALSE}
# Set some useful chunk options for code chunks.
knitr::opts_chunk$set(
  echo = TRUE,
  error = TRUE,    # display errors but keep going.
  comment = "",    # don't add '##' characters before output.
  message = FALSE  # don't include messages in the knitted output (check them in RStudio)
  )
library(tidyverse)
theme_set(theme_bw())
```

## GapMinder Country Populations

This homework explores this slightly altered, multi-table dataset of world populations.

## The Purpose

Our goal is to do an EDA (Exploratory Data Analysis) on world populations by country.

## Loading Data

```{r reading csv files}
country_codes <- read_csv ("data/country-codes.csv")
country_pop <- read_csv("data/country-population.csv")
```

## Querying Raw Population Data

```{r pivoting the data}
country_pop2 <- country_pop %>% 
  gather( key = "year", value = "pop", -country, convert = TRUE)

country_population <-  full_join(country_pop2, country_codes, by = c ("country" = "code"))
```


```{r facet wrap}
country_population %>%
  filter(country %in% c ("aus", "bgd", "chn","idn", "dom", "deu","kor", "usa", "gbr")) %>%
  filter(year %in% c("1970", "1971", "1972", "1973", "1974", "1975", "1976", "1977", "1978", "1979", "1980")) %>%
  ggplot() +
  aes(x = year, y = pop) +
  labs(title = "Populations over time", x = "Year", y = "Population")+
  facet_wrap(~country)+
  geom_line()
  
```

## Exploring the Data
How does some of the newly industrialized country's population change over time?
```{r comparing indonesia, malaysia, india}
country_population %>%
  filter(country %in% c ("idn", "ind", "mys")) %>%
  ggplot() +
  aes(x = year, y = pop) +
  labs(title = "Some Industrialized Countries Populations over time", x = "Year", y = "Population")+
  facet_wrap(~country)+
  geom_line()
```
It is interesting how 3 countries that are all newly industrialized have such difference in population growth. I honestly expected the graphs to look similar but as we can see, India has a really steep growth in population whereas Malaysia just has a little increase in population. Indonesia's population looks like it is growing faster than before, but still not as fast as India.


