---
title: "Hotel Booking Demand Predictive Analytics"
author: "Sherise Immanuela & Tanya Nyamwiwa"
date: "`r Sys.Date()`"
output:
  html_document:
    code_download: true
---

```{r setup, include=FALSE}
# Set some useful chunk options for code chunks.
knitr::opts_chunk$set(
  echo = TRUE,
  error = TRUE,    # display errors but keep going.
  comment = "",    # don't add '##' characters before output.
  message = FALSE  # don't include messages in the knitted output (check them in RStudio)
  )
library(tidyverse)
theme_set(theme_bw())
```

## Overview

### Question
What makes people cancel their hotel bookings?

### Summary of Findings

## Prior Work

### Summary and critique of prior work

Ideally this section compares two different prior analyses on similar topics.

### How this project extends prior work

## Approach

### Problem Description

### Data

#### Provenance

#### Structure

#### Appropriateness for task

### Modeling Question and Approach

## Exploratory Data Analysis

```{r}
hotel_bookings <- read.csv("data/hotel_bookings.csv")
```

Which months are the busiest?
```{r}
#hotel_bookings |>
#  ggplot() +
#  aes(x = arrival_date_month, y = adr) +
#  geom_line(aes(color = hotel))
  #scale_fill_viridis_d() +
  #theme_minimal()
```

```{r}
hotel_bookings$arrival_date_month = factor(hotel_bookings$arrival_date_month, levels = month.name)

ggplot(hotel_bookings, aes(x=arrival_date_month, fill = hotel))+
  geom_bar(position = 'dodge')+
  ylab('Number of Hotel Bookings ')+
  xlab('Month')+
  ggtitle('Average Number of Hotel Bookings per Month')+
  labs(fill='Hotel')+
  theme_minimal()

```

```{r}
checkout = hotel_bookings  %>% filter(reservation_status == 'Check-Out') 
dim(checkout)

ggplot(checkout, aes(customer_type, fill=hotel)) +
  geom_bar(stat='count', position=position_dodge())+
  ylab('Count')+
  xlab('Booking type')+
  ggtitle('Type of Booking by Different Hotel Type') +
  labs(fill='Hotel type')+
  theme_minimal()
```

```{r}
hotel_bookings |>
  ggplot() +
  aes(lead_time)+ 
  geom_histogram(binwidth = 0.8,col='turquoise' )+ 
  facet_wrap(~ is_canceled)
```

### Individual Variables
### Combinations of Variables
## Modeling
## Summary
### Discussion of Findings
This section must compare two different approaches (models, representation choices, evaluation methods, or whatever you have chosen).
### Limitations and Social / Ethical Considerations
### Future Directions
